---
title: "Final project of STAT184-Zhefeng He"
output: html_notebook
---

#I.Data Access

###Following are packages we gonna use in the project.
###Loading data collected from:
###http://www.ungei.org/resources/files/unesco_mexico124210eo.pdf This demonstrates the school attendence of Mexico
###http://www.electionresources.org/mx/deputies.php?election=2000&state=32
###election map of Mexico
###Loading x1:97 election map, x2:2000 election map, x3: school attandenceof 1990s
```{r}
library(readxl)
library(ggplot2)
library(httr)
library(dplyr)
library(gridExtra)
library(tidyverse)
library(repmis)

library(httr)

x3 <- read_excel("per.xlsx")
View(x3)

x2<-read_excel("20poll.xlsx")
View(x2)

x1<-read_excel("97poll.xlsx")

View(x1)
```
###Here WE inpsect the infomration of datasets and check variables.
```{r}
summary(x1)
summary(x2)
summary(x3)
str(x1)
str(x2)
str(x3)

```
###As there are special characters in election 20, so using regular expression to clear it.
###Also, in x1, the percentage number is expressed in"0.XXX", mutiple it by 100.
```{r}
x2 <-
  x2 %>%
  mutate(er20 = gsub(pattern = "[#,%]", replacement = "", er20)) %>%
  mutate(er20=as.numeric(er20))
x2

x1 <-
  x1 %>%
  mutate(er97=er97*100)
x1
```
# Data sets combination
###Here we join two tables for further analysis.
```{r}
me <-
  x1 %>%
  inner_join(x2, by=c("name")) %>%
  inner_join(x3, by=c("name"))
me
```
#Create the variable called "change" to evalute. 
```{r}
me2 <-
  me %>%
  mutate(change=(er20-er97))
me2
```

###Gather the me2 dataset to collect election result under same variable "name".
###Generate the graphs to show the distribution of elections data.
###As you can see a general increase on election 2000 compared to election 1997.
```{r} 
me1<-  
  me2 %>%
  group_by(name) %>%
  gather( key=election, value=votes, er97, er20 )
me1 %>%
  ggplot(aes(x=name, y=votes, color=election)) +
  geom_point()
me1 %>%
  ggplot(aes(x = name, y = votes)) +
  geom_point()+
  facet_wrap(~election)

```
### Here we measure central tendency and standard deviation of Independent variable with mean and sd.
### so the mean of school attendance is 6.264516% and deivation of positive & negative 2.608646 %.
```{r}
mean(me2$perceatend)
sd(me2$perceatend)
```
### Back to the me2 dataset.
### Rearrange the me2 in the rank of most votes changing state for further research
### checking the distribution of independent variables by states through the histogram graph.

```{r}
me2 <-
  me2 %>%
  filter(change >= 0) %>%
  group_by(name) %>%
  arrange(desc(change))
me2 %>%
  ggplot(aes(x = change, color=name)) +
  geom_histogram() 


```
###Using the me2 to generate the distribution of dependent variables by independent variables
### add density to see the changing trend.
###the density tells us a rough negative relationship.
```{r}
p1<-ggplot(me2,aes(x=perceatend,y=er20))+
  geom_point(size=2, color='red', alpha=0.9)+  #Better setting size and color
  geom_smooth(method='lm')+
  labs(x="Scholl attending rate by districts",
       y="MEXICO 2000 election polls of PAN (Largest opposite party)",
       title = "Mexico 2000 election votes for PAN in districts by school attending percentage in districts",
       caption = "The Educational and Social Situation in Mexico:
The Future of Girls’ Education，Interests, Institutions, and Reformers:
The Politics of Education Decentralization in Mexico, Election Resources on the Internet:
Federal Elections in Mexico
       ")
p1
```
###building a liner model to analyze the relationship
### Low R-square means the relationship is not significant.
```{r}
fg1<-lm(change~perceatend,data = me2)
summary(fg1)
```
###graph build based on the model support the result of low R-square.
```{r}
pg2<-ggplot(me2,aes(x=perceatend,y=change))+
  xlim(min(me2$perceatend),max(me2$perceatend))+
  ylim(min(me2$change),max(me2$change))+
  geom_abline(intercept =fg1$coefficients[1], #a+bx+e
              slope = fg1$coefficients[2],
              size=1,
              color="Blue")+
  labs(x="School attending by districts",
       y="districts votes change",
       title = "1997-2000 Mexico PAN support changes by districts",
       caption = "The Educational and Social Situation in Mexico:
         The Future of Girls’ Education，Interests, Institutions, and Reformers:
         The Politics of Education Decentralization in Mexico, Election Resources on the Internet:
         Federal Elections in Mexico")
pg2
```
###Why is not statitical signifcant?
###probably because the wide data distribution.
###the machine-learning graph verified the wide data distribution.
###This is we why cannt conclude a statical significant result.
```{r}
Mex <- dist(me2)
hc <- hclust(Mex)
plot(hc, hang=-1)

```